---
title: "Analysis"
author: "Sofie McComb"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Packages

```{r}
library(ggplot2)
library(tidyverse)
```


### Load Data

```{r}

load("Data/df/ticks_fulldata.Rda")
load("Data/df/mammals_fulldata.Rda")

```


### Nymph (Ixodes scap) and larvae (all species) time series for SERC and HARV site for 2018

Aggregate data by plot and collection date (or Month-Year) to calculate density of ticks by species/life stage. 

Need to use individual taxa counts over nymph counts, otherwise ignoring taxa relevant information (not only counting ixodes scapularis when subset). But need to use larvaCount over individual taxa counts, since need all taxa for larva and not all identified (would be missing a lot).

####Aggregate

```{r}


#Nymph (Ixodes scapularis) Have to use individualCount to use 
nymph_2018_ixosca<-ticks_fulldata %>% 
  filter(sexOrAge == 'Nymph',
         scientificName == 'Ixodes scapularis',
         Year=="2018") %>% 
  rename(NymphCount=individualCount_taxa)
nymph_count_2018_ixosca<-aggregate(NymphCount ~ siteID + plotID + Year +MonthYear + date, data=nymph_2018_ixosca, FUN=sum)

#Larva (All species)
larva_2018<-ticks_fulldata %>% 
  filter(Year=="2018") %>% 
    rename(LarvaCount=larvaCount) %>% 
  select(siteID, plotID, Year, MonthYear, date, LarvaCount) %>%  #only necessary columns before remove duplicates so count is accurate
  distinct()#remove duplicate rows (made for taxa ID/pathogens) so each observation only count once
larva_count_2018<-aggregate(LarvaCount ~ siteID + plotID + Year +MonthYear + date, data=larva_2018, FUN=sum) #Aggregate just remove 2 rows with NA value


#Combine dataframes with fulljoin
count_2018<-full_join(nymph_count_2018_ixosca, larva_count_2018, by=c("siteID", "plotID", "Year", "MonthYear", "date")) 



```

#### Visualize: SERC
```{r}

#All SERC plots together
SERC_count_2018_allplots<-count_2018 %>% 
  filter(siteID=="SERC") %>% 
  mutate(Date=substr(date, 0,10) %>% lubridate::ymd()) %>% 
    group_by(siteID, Year, MonthYear, Date) %>% 
  summarise(Nymph=sum(NymphCount, na.rm = TRUE),
            Larva=sum(LarvaCount, na.rm = TRUE)) %>% 
  gather(Stage, Count, Nymph:Larva) %>% 
  mutate(Stage=as.factor(Stage))
SERC_count_2018_allplots$Stage<-relevel(SERC_count_2018_allplots$Stage, "Nymph")#Relevel so nymph first in legend

SERC_allplots<-ggplot(SERC_count_2018_allplots, aes(x = Date, y=Count, group=Stage, color=Stage)) + 
  geom_line(size=1)+
  theme_minimal()+
  scale_color_manual(values=c("#00AFBB", "#E7B800"))

SERC_allplots


```

```{r}

#SERC plots as facet
SERC_count_2018<-count_2018 %>% 
  filter(siteID=="SERC") %>% 
  mutate(Date=substr(date, 0,10) %>% lubridate::ymd()) %>% 
  group_by(siteID, plotID, Year, MonthYear, Date) %>% 
  summarise(Nymph=sum(NymphCount, na.rm = TRUE),
            Larva=sum(LarvaCount, na.rm = TRUE)) %>% 
  gather(Stage, Count, Nymph:Larva) %>% 
  mutate(Stage=as.factor(Stage))
SERC_count_2018$Stage<-relevel(SERC_count_2018$Stage, "Nymph")#Relevel so nymph first in legend


SERC_facetplots<-ggplot(SERC_count_2018, aes(x = Date, y=Count, group=Stage, color=Stage)) + 
  geom_line(size=1)+
  theme_minimal()+
  scale_color_manual(values=c("#00AFBB", "#E7B800"))+
  facet_wrap(~plotID)


SERC_facetplots


```

#### Visualize: HARV

```{r}

#All HARV plots together
HARV_count_2018_allplots<-count_2018 %>% 
  filter(siteID=="HARV") %>% 
  mutate(Date=substr(date, 0,10) %>% lubridate::ymd()) %>% 
    group_by(siteID, Year, MonthYear, Date) %>% 
  summarise(Nymph=sum(NymphCount, na.rm = TRUE),
            Larva=sum(LarvaCount, na.rm = TRUE)) %>% 
  gather(Stage, Count, Nymph:Larva) %>% 
  mutate(Stage=as.factor(Stage))
HARV_count_2018_allplots$Stage<-relevel(HARV_count_2018_allplots$Stage, "Nymph")#Relevel so nymph first in legend

HARV_allplots<-ggplot(HARV_count_2018_allplots, aes(x = Date, y=Count, group=Stage, color=Stage)) + 
  geom_line(size=1)+
  theme_minimal()+
  scale_color_manual(values=c("#00AFBB", "#E7B800"))

HARV_allplots


```

```{r}

#HARV plots as facet
HARV_count_2018<-count_2018 %>% 
  filter(siteID=="HARV") %>% 
  mutate(Date=substr(date, 0,10) %>% lubridate::ymd()) %>% 
  group_by(siteID, plotID, Year, MonthYear, Date) %>% 
  summarise(Nymph=sum(NymphCount, na.rm = TRUE),
            Larva=sum(LarvaCount, na.rm = TRUE)) %>% 
  gather(Stage, Count, Nymph:Larva) %>% 
  mutate(Stage=as.factor(Stage))
HARV_count_2018$Stage<-relevel(HARV_count_2018$Stage, "Nymph")#Relevel so nymph first in legend


HARV_facetplots<-ggplot(HARV_count_2018, aes(x = Date, y=Count, group=Stage, color=Stage)) + 
  geom_line(size=1)+
  theme_minimal()+
  scale_color_manual(values=c("#00AFBB", "#E7B800"))+
  facet_wrap(~plotID)


HARV_facetplots


```






###Infection prevalence

Aggregate data by plot and collection date (or Month-Year) to calculate infection prevalence with different pathogens by tick species/plot/date.

[Pause until determine what infections, tick species, and life stages to aggregate around]



### Plot proportion of mammals trapped that had feeding ticks for each lifestage found at the SERC site across all years

#### Plot with # animals per trap in each plot/date

(e.g. 5 captures per 200 traps set in plot X on date Y = 0.025)

```{r}

mammals_df<-mammals_fulldata %>% 
  select(-nlcd, -lat, -long, -Month, -Day, -MonthName) %>% 
  mutate(Trap=ifelse(trapstatuscode == '2', 1,
                    ifelse(trapstatuscode == '3', 1,
                        ifelse(trapstatuscode== "4", 1,
                              ifelse(trapstatuscode=="5", 1,
                                     ifelse(trapstatuscode=="6",1,0)))))) #1 should be 0, as well as any others


#Aggregate to get # traps set in each plot-date combination
traps_set<-aggregate(Trap ~ siteID + plotID + Year +MonthYear + date, data=mammals_df, FUN=sum)

#Aggregate to get # captures in each plot-date combination
captures<-aggregate(capture~ siteID + plotID + Year +MonthYear + date, data=mammals_df, FUN=sum)

cap_trap<-full_join(captures, traps_set, by=c("siteID", "plotID", "Year", "MonthYear", "date")) %>% 
  mutate(cap_trap=capture/Trap) %>% 
  filter(!is.na(cap_trap))#Remove NA (where trap is 0)

serc_cap_trap<-cap_trap %>% filter(siteID=="SERC")

ggplot(serc_cap_trap, aes(x=date, y=cap_trap))+
  geom_line(size=1, color="#00AFBB")+
  facet_wrap(~plotID)+
  theme_minimal()+
  labs(x="Date", y="Trapping effort (# of Animals Captured/Traps Set)")


```


#### Plot # of animals with feeding ticks per total number of animals trapped in each plot/date 
(e.g. 2 animals with feeding ticks of 5 total animals captured)

```{r}

#Use mammals_df from above as base

mammals_ticks<-mammals_df %>% 
  mutate(larvae_Y=ifelse(larvalTicksAttached == 'Y', 1, 0),
         nymphs_Y=ifelse(nymphalTicksAttached == 'Y', 1, 0),
         adults_Y=ifelse(adultTicksAttached == 'Y', 1, 0))



#Aggregate number of animals with each type of feeding tick
larval_ticks<-aggregate(larvae_Y~siteID + plotID + Year +MonthYear + date, data=mammals_ticks, FUN=sum)
nymph_ticks<-aggregate(nymphs_Y~siteID + plotID + Year +MonthYear + date, data=mammals_ticks, FUN=sum)
adult_ticks<-aggregate(adults_Y~siteID + plotID + Year +MonthYear + date, data=mammals_ticks, FUN=sum)

#Join results together with cap_trap from previous section

mams_ticks<-cap_trap %>% 
  full_join(larval_ticks, by=c("siteID", "plotID", "Year", "MonthYear", "date")) %>% 
  full_join(nymph_ticks, by=c("siteID", "plotID", "Year", "MonthYear", "date")) %>% 
  full_join(adult_ticks, by=c("siteID", "plotID", "Year", "MonthYear", "date")) %>% 
  filter(!is.na(cap_trap)) %>% #where trap is 0
  mutate(larvae_cap=ifelse(capture==0, 0, larvae_Y/capture), #Calculate #animals with feeding ticks/#captured
         nymph_cap=ifelse(capture==0, 0, nymphs_Y/capture),
         adult_cap=ifelse(capture==0, 0, adults_Y/capture))


#Plot for SERC
serc_mams_ticks<-mams_ticks %>% filter(siteID=="SERC") %>% 
  rename(Larvae=larvae_cap, Nymph=nymph_cap, Adult=adult_cap) %>% 
  gather(stage, proportion, Larvae:Adult) %>% 
  mutate(Stage=as.factor(stage))
serc_mams_ticks$Stage<-relevel(serc_mams_ticks$Stage, "Nymph")#Relevel so nymph first in legend
serc_mams_ticks$Stage<-relevel(serc_mams_ticks$Stage, "Larvae")#Relevel so Larvae first in legend

ggplot(serc_mams_ticks, aes(x=date, y=proportion, group=Stage, color=Stage))+
  geom_line(size=1)+
  facet_wrap(~plotID)+
  theme_minimal()+
  labs(x="Date", y="Proportion Captured Animals with Feeding Ticks")+
  scale_color_manual(values=c("#E7B800","#00AFBB", "#90ee90"))


```



#### Plot density of animals with feeding ticks in each plot/date
(e.g. 2 animals with feeding ticks per 200 traps set in plot X on date Y).


```{r}

#Based off of mams_ticks from above

mams_traps<-mams_ticks %>% 
  mutate(larvae_traps=larvae_Y/Trap,
         nymph_traps=nymphs_Y/Trap,
         adult_traps=adults_Y/Trap)

#Plot for SERC
serc_mams_traps<-mams_traps %>% filter(siteID=="SERC") %>% 
  rename(Larvae=larvae_traps, Nymph=nymph_traps, Adult=adult_traps) %>% 
  gather(stage, proportion, Larvae:Adult) %>% 
  mutate(Stage=as.factor(stage))
serc_mams_traps$Stage<-relevel(serc_mams_traps$Stage, "Nymph")#Relevel so nymph first in legend
serc_mams_traps$Stage<-relevel(serc_mams_traps$Stage, "Larvae")#Relevel so Larvae first in legend

ggplot(serc_mams_traps, aes(x=date, y=proportion, group=Stage, color=Stage))+
  geom_line(size=1)+
  facet_wrap(~plotID)+
  theme_minimal()+
  labs(x="Date", y="Proportion Traps Set having Animals with Feeding Ticks")+
  scale_color_manual(values=c("#E7B800","#00AFBB", "#90ee90"))


```



