---
title: "Data Processing"
author: "Sofie McComb"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#NEON data exploration for Macroecosystems Proposal

###Load Packages

```{r packages}

rm(list=ls())

#devtools::install_github("NEONScience/NEON-geolocation/geoNEON")
library(neonUtilities)
library(ggplot2)
library(tidyverse)
library(geoNEON)
library(metScanR)


```


###Download and Stack Data

Only need to download and stack once (make sure data.table is updated)

```{r data}

#All sites:
# zipsByProduct(dpID="DP1.10093.001", site="all", package="basic", check.size=T) # ticks
# zipsByProduct(dpID="DP1.10092.001", site="all", package="basic", check.size=T) # tick pathogens
# zipsByProduct(dpID="DP1.10072.001", site="all", package="basic", check.size=T) # mammals

#Stack data:
# neonUtilities::stackByTable(filepath="filesToStack10093", folder=T)
# neonUtilities::stackByTable(filepath="filesToStack10072", folder=T)
# neonUtilities::stackByTable(filepath="filesToStack10092", folder=T)

```


### Read in data (not uploading to github)

```{r}

#Read in csv from stacked data: (need read.csv for logical columns)
#Tick observations
ticks_df <- read.csv("filesToStack10093/stackedFiles/tck_fielddata.csv")
#Ticks taxa
ticks_taxa_df <- read.csv("filesToStack10093/stackedFiles/tck_taxonomyProcessed.csv")
#Mammal data
mammals_df <- read.csv("filesToStack10072/stackedFiles/mam_pertrapnight.csv")
#Pathogens
tick_path_df <- read.csv("filesToStack10092/stackedFiles/tck_pathogen.csv")

```

### Data wrangle

####Clean-up each dataframe

```{r}

#Selecting needed columns
ticks<-ticks_df %>% 
  mutate(siteID=as.character(siteID),
         plotID=as.character(plotID),
         collectDate=as.character(collectDate),
         lat=decimalLatitude,
         long=decimalLongitude,
         nlcd=nlcdClass) %>% 
  select(siteID,  plotID, collectDate,
         targetTaxaPresent, adultCount,
         nymphCount, larvaCount,
         totalSampledArea, nlcd,
         lat, long
         )

ticks_taxa<-ticks_taxa_df %>% 
  mutate(siteID=substr(plotID, 1, 4),
         siteID=as.character(siteID),
         plotID=as.character(plotID),
         collectDate=as.character(collectDate),
         sciNameID=acceptedTaxonID,
         individualCount_taxa=individualCount) %>% 
  select(siteID, plotID, collectDate,
         scientificName,sciNameID,
         sexOrAge, individualCount_taxa
         )

tick_path<-tick_path_df %>% 
    mutate(siteID=as.character(siteID),
         plotID=as.character(plotID),
         collectDate=as.character(collectDate),
         lat=decimalLatitude,
         long=decimalLongitude,
         nlcd=nlcdClass,
         individualCount_path=individualCount) %>% 
  select(siteID, plotID, collectDate,
         testingID,individualCount_path,
         testResult,testPathogenName
         # nlcd, lat, long #otherwise combined twice
        )

mammals<-mammals_df %>% 
  # mutate_if(is.factor, as.character) %>%  #If want to switch factors to character columns
  mutate(siteID=as.character(siteID),
         plotID=as.character(plotID),
         collectDate=as.character(collectDate),
         lat=decimalLatitude,
         long=decimalLongitude,
         nlcd=nlcdClass,
         trapstatuscode=substr(trapStatus,0,1),
         capture=ifelse(trapstatuscode == '5', 1,
                    ifelse(trapstatuscode == '4', 1, 0))) %>%
  select(siteID, plotID, collectDate,
         trapstatuscode,trapStatus, capture,
         tagID, taxonID,
         scientificName, recapture,
         larvalTicksAttached, nymphalTicksAttached,
         adultTicksAttached,
         nlcd, lat, long
         )
         
         

```

#### Combine tick dataframes

```{r}

ticksdata<-full_join(ticks, ticks_taxa,
                      by=c("siteID", "plotID", "collectDate"))
ticks_data<-ticksdata %>% 
  full_join(tick_path, by=c("siteID", "plotID", "collectDate"))

```


#### Add Month and Year information to both datasets (ticks_data, mammals)

```{r}

ticks_fulldata<-ticks_data %>%   
  mutate(Year=substr(collectDate,0,4),
         Month=substr(collectDate,6,7),
         Day=substr(collectDate,9,10),
         Time=substr(collectDate,12,16)
         )%>% 
  mutate(MonthName=ifelse(Month=="01", "Jan",
                    ifelse(Month=="02", "Feb",
                     ifelse(Month=="03", "Mar",
                      ifelse(Month=="04", "Apr",
                        ifelse(Month=="05", "May",
                          ifelse(Month=="06", "Jun",
                            ifelse(Month=="07", "Jul",
                              ifelse(Month=="08", "Aug",
                               ifelse(Month=="09", "Sep",
                                ifelse(Month=="10", "Oct",
                                 ifelse(Month=="11", "Nov",
                                  ifelse(Month=="12", "Dec", NA))))))))))))) %>% 
  mutate(MonthYear=paste0(MonthName, "-", Year))



mammals_fulldata<-mammals %>% 
  mutate(Year=substr(collectDate,0,4),
         Month=substr(collectDate,6,7),
         Day=substr(collectDate,9,10)
         ) %>% 
  mutate(MonthName=ifelse(Month=="01", "Jan",
                    ifelse(Month=="02", "Feb",
                     ifelse(Month=="03", "Mar",
                      ifelse(Month=="04", "Apr",
                        ifelse(Month=="05", "May",
                          ifelse(Month=="06", "Jun",
                            ifelse(Month=="07", "Jul",
                              ifelse(Month=="08", "Aug",
                               ifelse(Month=="09", "Sep",
                                ifelse(Month=="10", "Oct",
                                 ifelse(Month=="11", "Nov",
                                  ifelse(Month=="12", "Dec", NA))))))))))))) %>% 
  mutate(MonthYear=paste0(MonthName, "-", Year))


```


#### Create versions to save that are only presence only data

```{r}

ticks_presence<-ticks_fulldata %>% 
  filter(targetTaxaPresent=="Y")

mammals_presence<-mammals_fulldata %>% 
  filter(capture==1)

```


#### Save datasets (not to github, too large)

```{r}

save(ticks_fulldata, file="Data/df/ticks_fulldata.Rda")
save(mammals_fulldata, file="Data/df/mammals_fulldata.Rda")
save(ticks_presence, file="Data/df/ticks_presence.Rda")
save(mammals_presence, file="Data/df/mammals_presence.Rda")

```


